---
title: "hw-06 data wrangling wrap up"
output:
   github_document
always_allow_html: yes
---

## Table of contents

-   [Question 1: Character Data](#question-1-character-data)
-   [Question 2: Writing Functions](#part-2-writing-functions)

## Question 1 - Character Data
Exercises from **Strings** chapter in **R for Data Science**.

## Initialize

* Load the required libraries:

```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(stringr))
suppressPackageStartupMessages(library(stringi))
```


### 14.2.5 String basics
**1. In code that doesn’t use stringr, you’ll often see paste() and paste0(). What’s the difference between the two functions? What stringr function are they equivalent to? How do the functions differ in their handling of NA?**

* `paste()` separates strings by spaces by default, whereas `paste0()` doesn't
* The stringr function `str_c()` is equivalent to `paste0()`
* In terms of how the functions handle NA: `str_c()` converts it to a missing value; paste0() converts it the string `"NA"`

**2. In your own words, describe the difference between the sep and collapse arguments to str_c().**

* `sep` is the string inserted between arguments
* `collapse` is the string that separates any elements of the character vector into a character vector of length one.

**3. Use str_length() and str_sub() to extract the middle character from a string. What will you do if the string has an even number of characters?**

We can select either the higher above the middle character using the *ceiling* function:

```{r}
x <- c("p", "pur", "purr", "purrr", "purrrr")
length <- str_length(x)
middle <- ceiling(length / 2)
str_sub(x, middle, middle)
```

**4. What does str_wrap() do? When might you want to use it?**

`str_wrap()` wraps text so that it fits within a certain width.

**5. What does str_trim() do? What’s the opposite of str_trim()?**

`str_trim()` trims the whitespace from a string. The opposite of `str_trim()` is `str_pad()`, which adds characters to each side.

```{r}
str_trim(" hello ", side = "left")
str_pad("hello", 5, side = "both")
str_pad("hello", 4, side = "right")
```

**6. Write a function that turns (e.g.) a vector c("a", "b", "c") into the string a, b, and c. Think carefully about what it should do if given a vector of length 0, 1, or 2.**

```{r}
str_convert <- function(x, delim = ",") {
  n <- length(x)
  if (n == 0) { #Tells the function how to handle length 0
    ""
  } else if (n == 1) { #Tells the function how to handle length 1
    x
  } else if (n == 2) { #Tells the function how to handle length 2
    # no comma before and when n == 2
    str_c(x[[1]], "and", x[[2]], sep = " ")
  } else { #Tells the function how to handle all other lengths
    not_last <- str_c(x[seq_len(n - 1)], delim)
    last <- str_c("and", x[[n]], sep = " ")
    str_c(c(not_last, last), collapse = " ")
  }
}

#Test function
str_convert(c("a", "b", "c", "d"))
```

### 14.3.1.1 Matching Patterns with Regex - Basic Matches

**1. Explain why each of these strings don’t match a "\: "\", "\\", "\\\"."**

-   `"\"`: Escapes the next character in the R string
-   `"\\"`: Resolves to `\` in the regular expression, which escapes the next character in the regular expression.
-   `"\\\"`: Resolves to a literal backslash in the regular expression, and escapes the next character.

**2. How would you match the sequence "'\"?**
```{}
str_view("\"'\\", "\"'\\\\")
```

**3. What patterns will the regular expression "\..\..\.."" match? How would you represent it as a string?**

```{r}
str_view(c(".a.b.c", ".a.b", "....."), c("\\..\\..\\.."))
```

### 14.3.2.1 Matching Patterns with Regex - Anchors

**1. How would you match the literal string "$^$"?**
```{r}
str_view(c("$^$", "ab$^$sfas"), "^\\$\\^\\$$")
```

**2. Given the corpus of common words in stringr::words, create regular expressions that find all words that:**

* Start with “y”
```{r}
    str_view(stringr::words, "^y", match =TRUE)
    ```

* End with “x”
```{r}
    str_view(stringr::words, "x$", match = TRUE)
    ```

* Are exactly three letters long.
```{r}
    str_view(stringr::words, "^...$", match = TRUE)
    ```

* Have seven letters or more.
```{r}
    str_view(stringr::words, ".......", match = TRUE)
    ```


### 14.3.3.1 Matching Patterns with Regex - Character classes and alternatives

**1. Create regular expressions to find all words that:**

* Start with a vowel
```{r}
    str_view(stringr::words, "^[aeiou]", match = TRUE)
    ```
    
* That only contain consonants. (Hint: thinking about matching “not”-vowels.)
```{r}
    str_view(stringr::words, "^[^aeiou]+$", match=TRUE)
    ```
    
* End with ed, but not with eed.
```{r}
    str_view(stringr::words, "^ed$|[^e]ed$", match = TRUE)
    ```
    
* End with ing or ise.
```{r}
    str_view(stringr::words, "i(ng|se)$", match = TRUE)
    ```


**2. Empirically verify the rule “i before e except after c”.**

```{r}
str_view(stringr::words, "(cei|[^c]ie)", match = TRUE)
str_view(stringr::words, "(cie|[^c]ei)", match = TRUE)
sum(str_detect(stringr::words, "(cei|[^c]ie)")) #count the number of words in each case
sum(str_detect(stringr::words, "(cie|[^c]ei)"))

```

**3. Is “q” always followed by a “u”?**

Yes, in the given database.

```{r}
str_view(stringr::words, "q[^u]", match = TRUE)
```

**4. Write a regular expression that matches a word if it’s probably written in British English, not American English.**

Assuming British English tends to use the following:

-   "ou" instead of "o"
-   ends in `ise` instead of `ize`

Use the regex `ou|ise$`.

**5. Create a regular expression that will match telephone numbers as commonly written in your country.**

```{r}
x <- c("123-456-7890", "1235-2351")
str_view(x, "\\d\\d\\d-\\d\\d\\d-\\d\\d\\d\\d")
```

### 14.3.4.1 Matching Patterns with Regex - Repetition
**1. Describe the equivalents of ?, +, * in {m,n} form.**

* `?` = `{0,1}` = Match at most 1
* `+` = `{1,}` = Match 1 or more
* `*` = `{0,}` = Match 0 or more

**2. Describe in words what these regular expressions match:**

* "^.*$""
Will match any string

* "\\{.+\\}"
Will match any string with curly braces surrounding at least one character

* "\d{4}-\d{2}-\d{2}""
Will match four digits followed by a hyphen, followed by two digits followed by a hyphen, followed by another two digits
     
* "\\\\{4}"
Will match four backslashes

**3. Create regular expressions to find all words that:**

* Start with three consonants.
```{r}
    str_view(words, "^[^aeiou]{3}")
    ```
    
* Have three or more vowels in a row.
```{r}
    str_view(words, "[aeiou]{3,}")
    ```
    
* Have two or more vowel-consonant pairs in a row.
```{r}
    str_view(words, "([aeiou][^aeiou]){2,}")
    ```

### 14.3.5.1 Matching Patterns with Regex - Grouping and backreferences

**1. Describe, in words, what these expressions will match:**
* "(.)\1\1""
The same character appearing three times in a row`

* "(.)(.)\\2\\1"
A pair of characters followed by the same pair of characters in reversed order

* "(..)\1""
Any two characters repeated.

* "(.).\\1.\\1"
A character followed by any character, the original character, any other character, the original character again

* "(.)(.)(.).*\\3\\2\\1"
Three characters followed by zero or more characters of any kind followed by the same three characters but in reverse order

**2. Construct regular expressions to match words that:**

* Start and end with the same character.
```{r}
    str_view(stringr::words, "^(.)((.*\\1$)|\\1?$)", match = TRUE)
    ```
    
* Contain a repeated pair of letters (e.g. “church” contains “ch” repeated twice.)
```{r}
    str_view(words, "(..).*\\1")
    ```
    
* Contain one letter repeated in at least three places (e.g. “eleven” contains three “e”s.)
```{r}
    str_subset(str_to_lower(words), "([a-z]).*\\1.*\\1")
    ```

### 14.4.2 Tools

**1. For each of the following challenges, try solving it by using both a single regular expression, and a combination of multiple str_detect() calls.**

* Find all words that start or end with x:
```{r}
    words[str_detect(words, "^x|x$")]
    start_with_x <- str_detect(words, "^x")
    end_with_x <- str_detect(words, "x$")
    words[start_with_x | end_with_x]
    ```

* Find all words that start with a vowel and end with a consonant:
```{r}
    str_subset(words, "^[aeiou].*[^aeiou]$") %>% head()
    start_with_vowel <- str_detect(words, "^[aeiou]")
    end_with_consonant <- str_detect(words, "[^aeiou]$")
    words[start_with_vowel & end_with_consonant] %>% head()
    ```

* Are there any words that contain at least one of each different vowel?

No
```{r}
    pattern <-
      cross(rerun(5, c("a", "e", "i", "o", "u")),
            .filter = function(...) {
              x <- as.character(unlist(list(...)))
              length(x) != length(unique(x))
            }) %>%
      map_chr(~ str_c(unlist(.x), collapse = ".*")) %>%
      str_c(collapse = "|")

    ```

**2. What word has the highest number of vowels? What word has the highest proportion of vowels? (Hint: what is the denominator?)**

Highest number of vowels:
```{r}
    vowels <- str_count(words, "[aeiou]")
    words[which(vowels == max(vowels))]
    ```

Highest proportion of vowels:
```{r}
    prop_vowels <- str_count(words, "[aeiou]") / str_length(words)
    words[which(prop_vowels == max(prop_vowels))]
    ```

### 14.4.3.1 Tools - Extract matches
**1. In the previous example, you might have noticed that the regular expression matched “flickered”, which is not a colour. Modify the regex to fix the problem.**

We want to only match colors in which the entire word is the name of the color, so we'll do this by adding a `\b` to indicate a word boundary:
```{r}
colours <- c("red", "orange", "yellow", "green", "blue", "purple")
colour_match <- str_c(colours, collapse = "|")
colour_match2 <- str_c("\\b(", str_c(colours, collapse = "|"), ")\\b")
colour_match2
```

**2. From the Harvard sentences data, extract:**

* The first word from each sentence.
```{r}
    str_extract(sentences, "[a-zA-Z]+") %>% head()
```

* All words ending in ing.
```{r}
    pattern <- "\\b[A-Za-z]+ing\\b"
    sentences_with_ing <- str_detect(sentences, pattern)
    unique(unlist(str_extract_all(sentences[sentences_with_ing], pattern))) %>%
      head()
```

* All plurals.
```{r}
    unique(unlist(str_extract_all(sentences, "\\b[A-Za-z]{3,}s\\b"))) %>%
      head()
```

### 14.4.4.1 Tools - Grouped matches

**1. Find all words that come after a “number” like “one”, “two”, “three” etc. Pull out both the number and the word.**
```{r}
numword <- "(one|two|three|four|five|six|seven|eight|nine|ten) +(\\S+)"
sentences[str_detect(sentences, numword)] %>%
  str_extract(numword)
```

**2. Find all contractions. Separate out the pieces before and after the apostrophe.**

```{r}
contraction <- "([A-Za-z]+)'([A-Za-z]+)"
sentences %>%
  `[`(str_detect(sentences, contraction)) %>%
  str_extract(contraction)
```

### 14.4.5.1 Tools - Replacing matches
**1. Replace all forward slashes in a string with backslashes.**
```{r}
backslashed <- str_replace_all("i/see/you", "\\/", "\\\\")
writeLines(backslashed)
```

**2. Implement a simple version of str_to_lower() using replace_all().**
```{r}
lower <- str_replace_all(words, c("A"="a", "B"="b", "C"="c", "D"="d", "E"="e", "F"="f", "G"="g", "H"="h", "I"="i", "J"="j", "K"="k", "L"="l", "M"="m", "N"="n", "O"="o", "P"="p", "Q"="q", "R"="r", "S"="s", "T"="t", "U"="u", "V"="v", "W"="w", "X"="x", "Y"="y", "Z"="z"))
```

**3. Switch the first and last letters in words. Which of those strings are still words?**

```{r}
swapped <- str_replace_all(words, "^([A-Za-z])(.*)([a-z])$", "\\3\\2\\1")
intersect(swapped,words)
```

### 14.4.6.1 Splitting
**1. Split up a string like "apples, pears, and bananas" into individual components.**
```{r}
x <- c("apples, pears, and bananas")
str_split(x, ", +(and +)?")[[1]]
```

**2. Why is it better to split up by boundary("word") than " "?**

`boundary("word")` recognizes non-space punctuation that splits words and removes punctuation while retaining internal non-letter characters that are parts of the word.

**3. What does splitting with an empty string ("") do?**

It splits the string into individual characters.
```{r}
str_split("ab. cd|agt", "")[[1]]
```

### 14.5.1 Other types of patterns
**1. How would you find all strings containing \ with regex() vs. with fixed()?**
```{r}
str_subset(c("a\\b", "ab"), "\\\\")
str_subset(c("a\\b", "ab"), fixed("\\"))
```

**2. What are the five most common words in sentences?**
```{r}
str_extract_all(sentences, boundary("word")) %>%
  unlist() %>%
  str_to_lower() %>%
  tibble() %>%
  set_names("word") %>%
  group_by(word) %>%
  count(sort = TRUE) %>%
  head(5)
```

### 14.7.1 stringi

**1. Find the stringi functions that:**

* Count the number of words.
`stri_count_words()`

* Find duplicated strings.
`stri_duplicated()`

* Generate random text.
`stri_rand_lipsum()`; `stri_rand_strings()`; `stri_rand_shuffle()`

**2. How do you control the language that stri_sort() uses for sorting?**
 
 `stri_sort(..., locale = ...)`

## Question 2 - Write a function

Write one (or more) functions that do something useful to pieces of the Gapminder or Singer data. It is logical to think about computing on the mini-data frames corresponding to the data for each specific country, location, year, band, album, … This would pair well with the prompt below about working with a nested data frame, as you could apply your function there.
 You could generalize that to do quadratic regression (include a squared term) or use robust regression, using MASS::rlm() or robustbase::lmrob()

### Initialize

* Load the required libraries:

```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(stringr))
suppressPackageStartupMessages(library(stringi))
```

### Method

* bla

### Code

```{r}
#sub_dat <- gapminder %>% 
 # filter(continent %in% c("Europe", "Asia")) %>% 
  #filter(year %in% c(2007, 2002))
#head(sub_dat)


#make_plot <- function(dat){
 # plt = ggplot(dat, aes(continent, gdpPercap, fill = continent)) + geom_boxplot() + facet_wrap(~year)
  #return(plt)
#}
#make_plot(sub_dat)
```




